
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
author: "Sophia Drezek"
date: "9-12-2022"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------
 \

1. Create the following data frame

| Rank 	| Age 	| Name  	|
|:------	|:-----	|:-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|


```{r}
df = data.frame(Rank = c(0:3),
                Age = c(28,34, 29, 42),
                Name = c('Tom', 'Jack', 'Steve', 'Ricky'))
df
```


2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).

```{r}
df <-  read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')

df
```


2. Show the names of the variables in the data

```{r}
names(df)
```


3. How many columns and rows the data have? 

```{r}
num_col = ncol(df)
num_col

num_row = nrow(df)
num_row
```


4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?

```{r}
sum(is.na(df))

colSums(is.na(df))

max(colSums(is.na(df)))

```


5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 

```{r}
class(df$date)
```

```{r}
df$date = as.Date(df$date)
class(df$date)
```


6. Capitalize the names of all the variables

```{r}
names(df) <- toupper(names(df))
names(df)
```


7. Find the average number of cases per day. Find the maximum cases a day.  

```{r}
avg_cases <- mean(df$CASES)
avg_cases

max_cases <- max(df$CASES)
max_cases

```


8. How many states are there in the data?  

```{r}
num_state <- unique(df$STATE)
length(num_state)
```


9. Create a new variable `weekdays` to store the weekday for each rows. 

```{r}
df$weekdays <- weekdays(df$DATE)
```


10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day

```{r}
df$DEATH2 <- ifelse(df$DEATHS==0, 
                        'no_death','has_death')
```

Find the frequency and relative frequency of `no_death` and `has_death`. 

```{r}
table(df$DEATH2)
```


11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)

```{r}
summary(df$DEATHS)
```


12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3

```{r}
library(tidyverse)
df$DEATH3 <- case_when(df$DEATHS<839 ~ 'low_death',
                       df$DEATHS<12680 ~ 'mid_death',
                       TRUE~'high_death')
table(df$DEATH3)
```

13. Find the average cases in Rhode Island in 2021

```{r}
df1 <- df[(df$DATE>='2021-01-01')&(df$STATE=='Rhode Island'),]
by(df1$CASES, df1$STATE, mean)

```


14. Find the median cases by weekdays in Rhode Island in 2021 

```{r}
df1 <- df[(df$DATE>='2021-01-01')&(df$STATE=='Rhode Island'),]
by(df1$CASES, df1$weekdays, median)
```


15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 

```{r}
library(lubridate)
df$MONTH <- month(df$DATE, label = TRUE)


df1 <- df[(df$DATE>='2021-01-01')&
            (df$STATE=='Rhode Island'),]

df2 <- df1[df1$MONTH %in% c('Jun','Jul','Aug','Sep'),]

by(df2$CASES, df2$MONTH, median)

```


16.  Find your own dataset, import it and implement the following functions on the data

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when

```{r}
setwd("C:/Users/student/OneDrive/OneDrive - Bryant University/Senior Semester 1/MATH 421/Assignments/Math 421")
df <- read.csv('titanic.csv')

head(df)

```

```{r}
str(df)

```

```{r}
names(df)
```


```{r}
mean(df$Age, na.rm=TRUE)
max(df$Age, na.rm=TRUE)
min(df$Age, na.rm=TRUE)
```

```{r}
table(df$SibSp)
```

```{r}
sum(is.na(df))
```

```{r}
colSums(is.na(df))
```

```{r}
class(df$Sex)

```


```{r}
cor(df$Pclass, df$Survived)
```

```{r}
by(df$Survived, df$Sex, FUN = mean)
```


```{r}
df$Survived2 <- ifelse(df$Survived==0, 
                        'Did not survive','Survived')
df
```


```{r}
library(dplyr)
df$Survived2 <- case_when(df$Survived==0 ~ 'Did not survive', 
                            TRUE ~ 'Survived')

df
```


17.  In the dataset in #16, practice the follows.  You can reuse the code of 16. 

- Create a categorical variable from a continuous variable
- From one categorical variable, create a new categorical variable with fewer categories


```{r}
df$Survived2 <- ifelse(df$Survived==0, 
                        'Did not survive','Survived')
df
```


```{r}
df$Age2 <- case_when(df$Age<12 ~ 'child',
                     df$Age<20 ~ 'teenager',
                     df$Age<50 ~ 'middle aged',
                     df$Age<70 ~ 'older adults',
                     TRUE~'elderly')
table(df$Age2)

```

```{r}
df$Age3 <- case_when(df$Age2 %in% c('child','teenager') ~ 'young',
                     df$Age2 %in% c('middle aged') ~ 'middle aged',
                     TRUE~'old')

table(df$Age3)
```



If you do not have a data, you can use [the titanic dataset](https://www.kaggle.com/competitions/titanic/overview), which can be downloaded at [this link](../data/titanic.csv)
