
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 8: Visualization - Making a bar race"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment8.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Install two packages `gganimate` and `gifski` then restart Rstudio. Using the `Adult Census Income` data, make an animation using `geom_point` and `transition_states`.

```{r}
library(gganimate)
library(gifski)

library(ggplot2)
library(tidyverse)
library(knitr)
```

```{r}
setwd("C:/Users/student/OneDrive/OneDrive - Bryant University/Senior Semester 1/MATH 421/Assignments/Math 421")
df <- read.csv('adult_census.csv')

head(df)
```

```{r}
df %>% ggplot(aes(x = hours.per.week,
           y = age, color=sex))+
  geom_point()+
  transition_states(marital.status)+
  labs(title = 'Marital Status: {closest_state}')
```


2. Using the `Adult Census Income` data, make an animation using `geom_bar` and `transition_states`.

```{r}
df %>% ggplot(aes(x = marital.status,
                  fill=race))+
  geom_bar(position = 'fill')+
  transition_states(sex)+
  labs(title = 'Sex: {closest_state}')
```


3. Use the WHO's dataset at [this link](https://covid19.who.int/WHO-COVID-19-global-data.csv). Make a top-10 bar race by months between countries on the number of deaths by Covid 19 in 2021.

```{r}
df <-  read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')

head(df)
```

```{r}
library(lubridate)

df$month = month(df$Date_reported)
```


```{r}
d1 <- df %>% 
  filter(Date_reported>='2021-01-01', Date_reported<'2022-01-01') %>% 
  group_by(month, Country) %>% 
  summarise(sum = sum(Cumulative_deaths))
d2 <- d1 %>% group_by(month) %>% mutate(rank=rank(-sum)) 
d3 <- d2 %>% filter(rank <= 10)
```


```{r}
a1 <- d3 %>% ggplot(aes(x=rank, y=sum, group=Country, fill=Country, label=Country)) + geom_col()+
    geom_text(aes(y = sum, label = Country), hjust = 1.4)+ 
    coord_flip(clip = "off", expand = FALSE) +scale_x_reverse()+
    labs(title = 'Month {closest_state}', x='', y='Total Number of Covid Deaths', fill='Country')+
    theme(plot.title = element_text(hjust = 1, size = 22),
          axis.ticks.y = element_blank(),
          axis.text.y  = element_blank()) + 
    transition_states(month)+
    ease_aes("cubic-in-out")
animate(a1, nframes = 400)
```


4. Make a bar race using a dataset of your own interest. You may use the dataset that we use in class (https://covidtracking.com/data/download/all-states-history.csv) but you should make a different bar race from ones in the slides. 

```{r}
setwd("C:/Users/student/OneDrive/OneDrive - Bryant University/Senior Semester 1/MATH 421/Assignments/Math 421")

df <- read_csv('hotel_bookings.csv')

head(df)
```


```{r}

d1 <- df %>% group_by(arrival_date_month, country) %>% summarise(mean = mean(lead_time))
d2 <- d1 %>% group_by(arrival_date_month) %>% mutate(rank=rank(-mean)) 
d3 <- d2 %>% filter(rank <= 10)
a1 <- d3 %>% ggplot(aes(x=rank, y=mean, group=country, fill=country, label=country)) + geom_col()+
    geom_text(aes(y = mean, label = country), hjust = 1.4)+ 
    coord_flip(clip = "off", expand = FALSE) +scale_x_reverse()+
    labs(title = 'Month {closest_state}', x='', y='Average lead_time', fill='country')+
    theme(plot.title = element_text(hjust = 1, size = 22),
          axis.ticks.y = element_blank(),
          axis.text.y  = element_blank()) + 
    transition_states(arrival_date_month)+
    ease_aes("cubic-in-out")
animate(a1, nframes = 100)
```


