
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 14: Twitters Mining with rtweet"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2021_assignment14.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


[Sample Codes](https://bryantstats.github.io/math421/slides/16_text_mining_rtweet.html)

-------

1. Pick a keyword or hashtag. Download the data associated with the keyword/hashtag. Plot at least 5 plots to visualize the data associated with the keyword/hashtag. All plots should have titles and captions. 

```{r}
library(rtweet) 
library(tidytext)
library(ggpubr) 
library(tidyverse) 
library(knitr)
library(lubridate)
```


```{r}
keyword_search = '#Nike'

df <- search_tweets(q = keyword_search, 
                        n = Inf, # number of tweets
                        include_rts = FALSE,
                        `-filter` = "replies",
                        lang = "en") %>% 
  mutate(created_at = ymd_hms(format(created_at, tz = "US/Eastern")))

write_csv(df, 'twitter_data.csv')
```

```{r}
df <- read_csv('twitter_data.csv')

head(df)
```

```{r}
ts_plot(df, "hours") +
  labs(x = NULL, y = NULL,
       title = "Frequency of Tweets by Time",
       subtitle = paste0(format(min(df$created_at), "%d %B %Y"), " to ", format(max(df$created_at),"%d %B %Y")),
       caption = "The frequency of tweets flucuates a lot over the time period.") +
  theme_minimal()
```


```{r}
df %>% 
  unnest_tokens(mentions, text, "tweets", to_lower = FALSE) %>%
  filter(str_detect(mentions, "^@")) %>%  
  count(mentions, sort = TRUE) %>%
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(mentions, n)))+
  geom_col()+
  labs(x = 'Number of mentions', y = 'Account Mentioned', title = 'Top 10 mentions in the tweets', caption='Poshmark is the account that is mentioned the most in the tweets.')
```

```{r}
library(RColorBrewer)

pal <- brewer.pal(8,"Dark2")
library(wordcloud) 

df %>% 
  filter(created_at>='2022-11-14') %>% 
  unnest_tokens(output = hashtag, input = text, token = "tweets") %>%
  filter(!hashtag %in% c("#cr7",'#ronaldo','#cristianoronaldo','#cristiano', '#mufc',"#manutd"), str_detect(hashtag, "^#")) %>% 
  count(hashtag, sort = TRUE) %>%
  mutate(hashtag = str_remove(hashtag, '#')) %>% 
  filter(hashtag != 'mufc') %>% 
  with(wordcloud(hashtag, n, random.order = FALSE, max.words = 50, colors = pal))
```


```{r}
df %>% 
  filter(created_at>='2022-11-14') %>% 
  unnest_tokens(output = hashtag, input = text, token = "tweets") %>%
  filter(!hashtag %in% c("#cr7",'#ronaldo','#cristianoronaldo','#cristiano', '#mufc',"#manutd"), str_detect(hashtag, "^#")) %>% 
  count(hashtag, sort = TRUE) %>%
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(hashtag, n)))+
  geom_col()+
  labs(x = 'Frequency', y = 'Hashtag', title = 'Top 10 hashtags in the tweets', caption = 'Nike, fashion, style, shopmycloset, and poshmark are the most common.')
```

```{r}
df %>% 
  ggplot(aes(x = retweet_count, y = favorite_count))+
  geom_point()+
  labs(x = 'Number of retweets', y = 'Number of Favorites', title = 'Number of favorites vs retweets', caption='The general trend is that if a tweet has more retweets, it has more favorites.')
```

```{r}
df %>%
    unnest_tokens(input = text, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(word, sort = TRUE) %>%
    inner_join(get_sentiments("nrc")) %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(y= reorder(sentiment, n), x = n))+geom_col()+
    labs(y='Relative Frequency', x ='', title='Sentiment of the Tweets', caption ='The tweets have more of a postive sentiment than a negative one.')
```

2. Choose a location then pick a trending keyword/hashtag in the location. Download the data associated with the keyword/hashtag. Plot at least 5 plots to visualize the data associated with the keyword/hashtag. All plots should have titles and captions. 


```{r}
trends_available()
```

```{r}
get_trends('Montreal')
```


```{r}
keyword_search = '#RIPTwitter'

df <- search_tweets(q = keyword_search, 
                        n = Inf, # number of tweets
                        include_rts = FALSE,
                        `-filter` = "replies",
                        lang = "en") %>% 
  mutate(created_at = ymd_hms(format(created_at, tz = "US/Eastern")))

write_csv(df, 'twitter_data1.csv')
```

```{r}
df <- read_csv('twitter_data1.csv')

head(df)
```


```{r}
ts_plot(df, "hours") +
  labs(x = NULL, y = NULL,
       title = "Frequency of Tweets by Time",
       subtitle = paste0(format(min(df$created_at), "%d %B %Y"), " to ", format(max(df$created_at),"%d %B %Y")),
       caption = "The frequency of tweets has one major spike between Nov. 17th and Nov. 18th.") +
  theme_minimal()
```

```{r}
df %>% 
  unnest_tokens(mentions, text, "tweets", to_lower = FALSE) %>%
  filter(str_detect(mentions, "^@")) %>%  
  count(mentions, sort = TRUE) %>%
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(mentions, n)))+
  geom_col()+
  labs(x = 'Number of mentions', y = 'Account Mentioned', title = 'Top 10 mentions in the tweets', caption='@elonmusk is the account that is mentioned the most in the tweets.')
```

```{r}
df %>% 
  filter(created_at>='2022-11-14') %>% 
  unnest_tokens(output = hashtag, input = text, token = "tweets") %>%
  filter(!hashtag %in% c("#cr7",'#ronaldo','#cristianoronaldo','#cristiano', '#mufc',"#manutd"), str_detect(hashtag, "^#")) %>% 
  count(hashtag, sort = TRUE) %>%
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(hashtag, n)))+
  geom_col()+
  labs(x = 'Frequency', y = 'Hashtag', title = 'Top 10 hashtags in the tweets', caption = '#riptwitter is the most common hashtag in these tweets.')
```

```{r}
df %>% 
  ggplot(aes(x = retweet_count, y = favorite_count))+
  geom_point()+
  labs(x = 'Number of retweets', y = 'Number of Favorites', title = 'Number of favorites vs retweets', caption='The general trend is that if a tweet has more retweets, it has more favorites.')
```

```{r}
df %>%
    unnest_tokens(input = text, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(word, sort = TRUE) %>%
    inner_join(get_sentiments("nrc")) %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(y= reorder(sentiment, n), x = n))+geom_col()+
    labs(y='Relative Frequency', x ='', title = "Sentiment of the Tweets", caption="There are approximately the same amount of negative sentiment and positive sentiment in the tweets.")
```

