
---
output: # rmarkdown::github_document
  html_document:
    theme: cerulean
    toc: true
  pdf_document: default
  word_document: default
title: "Assignment 5: Handle Missing Values"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment5.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Github Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**The data:** 

This assignment works with the Adult Census Data that can be downloaded at [this link.](https://bryantstats.github.io/math421/data/adult_census_missing.csv)

-------

**Questions**

1. Use `read_csv` to import the data.  Show the number of NAs for each columns.   

```{r}
library(tidyverse)
df <-  read_csv('https://bryantstats.github.io/math421/data/adult_census_missing.csv')

head(df)
```


```{r}
colSums(is.na(df))
```


2. Using function `aggr` the VIM package to plot the number of NAs for each column. 

```{r}
library(VIM)

aggr(df)
```


3. Find other forms of missing values. Hint: You can use the `table` function to check if there are suspicious categories (Unknown, for example) in a variable. What all other forms of missing values shown in the data. 

```{r}
table(df$age)
table(df$workclass)
table(df$education)
table(df$education.num)
table(df$marital.status)
table(df$occupation)
table(df$relationship)
table(df$race)
table(df$sex)
table(df$capital.gain)
table(df$capital.loss)
table(df$hours.per.week)
table(df$native.country)
table(df$income)
```


4. Replace all the forms of missing values found to NA

```{r}
df <- df %>% na_if('Unknown') %>% na_if('Not Known') %>% na_if('?')

colSums(is.na(df))
```


5. Replot the number of NAs for each column. 

```{r}
library(VIM)

aggr(df)
```


6. Approach 1 to handle NAs: remove all rows that have any NAs. Save the dataset after removing as a different data. The original data is unchanged (still have NAs).  How many rows left after removing?

```{r}
df1 <- drop_na(df)

nrow(df1)
```

7. Approach 2 to handle NAs:  Fill all the NAs with the previous or next value. (Hint: Using `fill` function).  Save the dataset after filling as a different data. The original data is unchanged (still have NAs). 

```{r}
df2 <- df %>% fill(age, workclass, education, marital.status,
                   occupation, relationship, sex, capital.gain,
                   native.country, .direction = 'updown')
colSums(is.na(df2))

```


8. Approach 3 to handle NAs: For numeric variable, replace the NAs by the median.  For categorical variables, replace the NAs to the majority. 

```{r}
df3 <- df

## numeric variables

median_age <- median(df$age, na.rm=TRUE)
df3$age <- replace_na(df$age, median_age)

median_cap_gain <- median(df$capital.gain, na.rm=TRUE)
df3$capital.gain <- replace_na(df$capital.gain, median_cap_gain)


## categorical variables

majority_workclass <- names(which.max(table(df$workclass)))
df3$workclass <- replace_na(df$workclass, majority_workclass)

majority_education <- names(which.max(table(df$education)))
df3$education <- replace_na(df$education, majority_education)

majority_marital.status <- names(which.max(table(df$marital.status)))
df3$marital.status <- replace_na(df$marital.status, majority_marital.status)

majority_occupation <- names(which.max(table(df$occupation)))
df3$occupation <- replace_na(df$occupation, majority_occupation)

majority_relationship <- names(which.max(table(df$relationship)))
df3$relationship <- replace_na(df$relationship, majority_relationship)

majority_sex <- names(which.max(table(df$sex)))
df3$sex <- replace_na(df$sex, majority_sex)

majority_native.country <- names(which.max(table(df$native.country)))
df3$native.country <- replace_na(df$native.country, majority_native.country)

colSums(is.na(df3))
```


