
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 7 - Extra Credits: Practice dplyr and ggplot2"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment7_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas



```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


-------

1. Import the data at https://bryantstats.github.io/math421/data/ds_salaries.csv

    The data description can be found at https://www.kaggle.com/datasets/saurabhshahane/data-science-jobs-salaries
    
```{r}
library(tidyverse)
df <-  read_csv('https://bryantstats.github.io/math421/data/ds_salaries.csv')

head(df)
```


2. Is there any column that contains NAs?

```{r}
colSums(is.na(df))
```


3. What is the most frequent job title in the data?  

```{r}
names(which.max(table(df$job_title)))
```


4. Find the average salary data of the three most frequent job titles in the data. 

```{r}
df %>% count(job_title) %>% arrange(-n) 
```

```{r}
df %>% 
  filter(job_title %in% c('Data Scientist', 'Data Engineer', 'Data Analyst')) %>% summarize(avg_salary = mean(salary_in_usd))
```


5. What is the job title that have the highest average salary?

```{r}
df %>% 
  group_by(job_title) %>% summarize(avg_salary = mean(salary_in_usd)) %>% arrange(-avg_salary)
```

6. Which country gives the highest salary on average? 

```{r}
df %>% 
  group_by(company_location) %>% summarize(avg_salary = mean(salary_in_usd)) %>% arrange(-avg_salary)
```


7. Plot five meaningful plots. All plots should have title, caption, and appropriate labels.

```{r}
df %>%
  filter(job_title %in% c('Data Scientist', 'Data Engineer', 'Data Analyst', 'Machine Learning Engineer', 'Research Scientist')) %>% 
  group_by(job_title) %>% 
  summarize(avg_salary = mean(salary_in_usd)) %>%
  ggplot()+ 
  geom_col(mapping=aes(x=job_title, y=avg_salary))+
  labs(x='Job Title', 
       y = 'Average Salary',
       title = 'Average Salary per Job Title', 
       caption = 'This shows the average salary for the top 5 job titles with the highest average salary in USD.')
```


```{r}
df %>%
  ggplot()+ 
  geom_point(mapping=aes(x=remote_ratio, y=salary_in_usd))+
  labs(x='Remote Ratio', 
       y = 'Salary',
       title = 'Remote Ratio vs. Salary', 
       caption = 'This shows the comparison between the ratio of remote workers and salary for each company')
```


```{r}
df %>%
  group_by(company_size) %>% 
  summarize(avg_salary = mean(salary_in_usd)) %>%
  ggplot()+ 
  geom_col(mapping=aes(x=company_size, y=avg_salary))+
  labs(x='Company Size', 
       y = 'Average Salary',
       title = 'Average Salary by Company Size', 
       caption = 'This shows the average salary by company size.')
```

```{r}
df %>%
  group_by(company_size, experience_level) %>% 
  summarize(avg_salary = mean(salary_in_usd)) %>%
  ggplot()+ 
  geom_col(mapping=aes(x=company_size, y=avg_salary, fill=experience_level))+
  labs(x='Company Size', 
       y = 'Average Salary',
       color = 'Experience Level',
       title = 'Average Salary by Company Size and Experience Level', 
       caption = 'This shows the average salary by company size and experience level.')
```

```{r}
df %>%   
  filter(job_title %in% c('Data Scientist', 'Data Engineer', 'Data Analyst', 'Machine Learning Engineer', 'Research Scientist')) %>% ggplot()+ 
  geom_point(mapping=aes(x=job_title, y=salary_in_usd))+
  facet_wrap(~experience_level)+
  labs(x='Job Title', 
       y = 'Salary in USD',
       title = 'Salary Differences Based on Experience Level', 
       caption = 'This shows the salary in USD for the top 5 jobs with the highest average salary broken down by each level of job experience.')
```

