
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 12: Predictive Modeling - Part 3"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment12.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


1. Use the `PimaIndiansDiabetes` dataset. Use 15% data for testing. Use cross-validation with of 7 folds to tune random forest `(method='ranger')`.  What are the parameters that produce the greatest accuracy? What is the testing accuracy. 

```{r}
library(mlbench)
library(tidyverse)
data(PimaIndiansDiabetes)
df <- tibble(PimaIndiansDiabetes)
head(df)
```

```{r}
df <- df %>% rename(target=diabetes)
```

```{r}
colSums(is.na(df))
```

```{r}
library(caret)
splitIndex <- createDataPartition(df$target, p = .75, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

```{r}
trControl = trainControl(method = "cv",
                         number = 7)

tuneGrid = expand.grid(mtry = 2:4,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(1:10))

forest_ranger <- train(target~., data=df_train, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
```

```{r}
plot(forest_ranger)
```

```{r}
forest_ranger
```

```{r}
pred <- predict(forest_ranger, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```


2. Use the `PimaIndiansDiabetes` dataset. Go to https://topepo.github.io/caret/available-models.html and pick a classification model.  Tune the classification model using cross-validation of 7 folds. 

```{r}
data(PimaIndiansDiabetes)
df <- tibble(PimaIndiansDiabetes)
```

```{r}
df <- df %>% rename(target=diabetes)
```

```{r}
splitIndex <- createDataPartition(df$target, p = .75, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

```{r}
trControl = trainControl(method = "cv",
                         number = 7)

tuneGrid = expand.grid(predFixed = 2:5,
                       minNode = 10)

forest_Rborist <- train(target~., data=df_train, 
                    method = "Rborist", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
```

```{r}
plot(forest_Rborist)
```

```{r}
forest_Rborist
```

```{r}
pred <- predict(forest_Rborist, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

3. (Model Comparison) Use the `PimaIndiansDiabetes` dataset. Pick two models at [this link](https://topepo.github.io/caret/available-models.html) to compare using 7-fold cross validation method. Evaluate the accuracy of the final model on the test data. 

```{r}
data(PimaIndiansDiabetes)
df <- tibble(PimaIndiansDiabetes)
```

```{r}
df <- df %>% rename(target=diabetes)
```

```{r}
splitIndex <- createDataPartition(df$target, p = .75, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

```{r}
trControl = trainControl(method = "cv",
                         number = 2)
```

```{r}
forest_ordinalRF <- train(target~., data=df_train, 
                    method = "ordinalRF", 
                                trControl = trControl)
```

```{r}
trControl = trainControl(method = "cv",
                         number = 7)

tree <- train(target~., data=df_train, 
                                method = "rpart2", 
                                trControl = trControl)
forest_ordinalRF <- train(target~., data=df_train, 
                    method = "ordinalRF", 
                                trControl = trControl)
lda <- train(target~., data=df_train, 
                                method = "lda", 
                                trControl = trControl)
results <- resamples(list('Decision Tree' = tree,
                          'Random Forest' = forest_ordinalRF,
                          'LDA'= lda))
bwplot(results)
```

```{r}
pred <- predict(forest_ordinalRF, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

