---
title: "Assignment3 Extra Credit"
author: "Sophia Drezek"
date: "2022-09-20"
output: html_document
---


16.  Find your own dataset, import it and implement the following functions on the data

- head
- str
- names
- mean, min, max
- table
- is.na
- colSums
- class
- cor
- by
- ifelse
- case_when

```{r}
setwd("C:/Users/student/OneDrive/OneDrive - Bryant University/Senior Semester 1/MATH 421/Assignments/Math 421")
df <- read.csv('adult_census.csv')

head(df)

```



```{r}
str(df)

```

```{r}
names(df)
```


```{r}
mean(df$age, na.rm=TRUE)
max(df$age, na.rm=TRUE)
min(df$age, na.rm=TRUE)
```

```{r}
table(df$race)
```

```{r}
sum(is.na(df))
```

```{r}
colSums(is.na(df))
```

```{r}
class(df$occupation)
```


```{r}
cor(df$education.num, df$hours.per.week)
```

```{r}
by(df$education.num, df$sex, FUN = mean)
```


```{r}
df$sex2 <- ifelse(df$sex=='Female', 
                        0,1)
head(df)
```


```{r}
library(dplyr)
df$age2 <- case_when(df$age<25 ~ 'Young adult', 
                     df$age<45 ~ 'Adult', 
                     df$age<65 ~ 'Middle-aged', 
                     df$age<80 ~ 'Senior Citizen', 
                     TRUE ~ 'Elderly')

head(df)
```



17.  In the dataset in #16, practice the follows.  You can reuse the code of 16. 

- Create a categorical variable from a continuous variable
- From one categorical variable, create a new categorical variable with fewer categories


```{r}
df$capital.loss2 <- case_when(df$capital.loss<1500 ~ 'low loss',
                              df$capital.loss<3000 ~ 'mid loss',
                              TRUE~ 'high loss')

table(df$capital.loss2)
```



```{r}
df$age3 <- case_when(df$age2 %in% c('Young adult','Adult') ~ 'Young',
                     df$age2 %in% c('Middle-aged') ~ 'Middle-aged',
                     TRUE~'old')

table(df$age3)
```


