
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 10: Predictive Modeling - Part 1"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment10.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Use the `Adult Census Income` dataset.  We will predict the income (whether or not it is more than 50k or not) of an adult. Import the dataset.  Partition the data into 80% training and 20% testing.  

```{r}
library(tidyverse)
setwd("C:/Users/student/OneDrive/OneDrive - Bryant University/Senior Semester 1/MATH 421/Assignments/Math 421")
df <- read_csv('adult_census.csv')

head(df)
```

```{r}
df <- df %>% rename(target=income)
```

```{r}
df <- df %>% 
  mutate(target = as.factor(target),
         workclass = as.factor(workclass),
         education = as.factor(education),
         education.num = as.factor(education.num),
         marital.status = as.factor(marital.status),
         occupation = as.factor(occupation),
         relationship = as.factor(relationship),
         race = as.factor(race),
         sex = as.factor(sex),
         native.country = as.factor(native.country)
         )
```


```{r}
library(caret)

set.seed(2020)
splitIndex <- createDataPartition(df$target, p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```


2. Practice Decision Tree.  Do the follows:

  - Use `rpart` package, create a decision tree with maximum depth of 3. 
  
```{r}
library(rpart)
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))
```
  
  - Calculate the accuracy of the model on the testing data. Notice that the positive outcome here is not `1` but `>50K` or `<50K`. 
  
```{r}
pred <- predict(tree_model, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

  - Plot the tree
```{r}
library(rattle)
fancyRpartPlot(tree_model)
```

  - Plot the variable importance by the tree
  
```{r}
tree_model$variable.importance
```
  
  
3. Create 3 more trees and compare the testing accuracy of these trees, which tree give the highest testing accuracy.

```{r}
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 2))
```

```{r}
pred <- predict(tree_model, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 8))
```

```{r}
pred <- predict(tree_model, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 1))
```

```{r}
pred <- predict(tree_model, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

4. Practice Random Forest.  Do the follows: 

  - Use `randomForest` package, create a random forest of 1000 trees. 
  
```{r}
library(randomForest)
forest_model = randomForest(target ~ ., data=df_train, ntree = 1000)
```

  - Calculate the accuracy of the model on the testing data. 

```{r}
pred <- predict(forest_model, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

  
  - Plot the variable importance by the forest
  
```{r}
importance(forest_model)
```

5. Create 3 more forests and compare the testing accuracy of these forests, which forest give the highest testing accuracy.

```{r}
forest_model = randomForest(target ~ ., data=df_train, ntree = 500)
```
```{r}
pred <- predict(forest_model, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
forest_model = randomForest(target ~ ., data=df_train, ntree = 1500)
```
```{r}
pred <- predict(forest_model, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
forest_model = randomForest(target ~ ., data=df_train, ntree = 100)
```
```{r}
pred <- predict(forest_model, df_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

6. What is the best model (in term of testing accuracy) among all models (including trees and forests) you have trained?

-Random Forest with 1500 trees has the highest accuracy!

