
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 13: Text Mining"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment13.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


[Sample Codes](text_mining_sample_codes2.html)

-------

### Netflix Data

**1.** Download the `netflix_titles` at this [link](../data/netflix_titles.csv).  Create a century variable taking two values:

    - '21' if the released_year is greater or equal to 2000, and
    
    - '20' otherwise. 
    
```{r}
library(tidyverse)
library(dplyr)
library(tidytext)
library(knitr)
```


```{r}
df <- read_csv('netflix_titles.csv')

head(df)
```

```{r}
df$century <- case_when(df$release_year >= 2000 ~ '21',
                       TRUE ~ '20')
```

    
**2. Word Frequency**    

  a. Convert the description to tokens, remove all the stop words. What are the top 10 frequent words of movies/TV Shows in the 20th century.  Plot the bar chart of the frequency of these words. 
  
```{r}
df %>% select(title, description) %>% 
  unnest_tokens(input = description, output = word) %>%
  anti_join(get_stopwords()) %>% 
  head(20)
```


```{r}
df %>% 
  filter(century=='20') %>% 
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(type, word, sort = TRUE) %>% 
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(word, n))) +
  geom_col() +
  labs(y = '', x = 'Frequency')
```

  b. What are the top 10 frequent words of movies/TV Shows in the 21st century. Plot the bar chart of the frequency of these words. Plot a side-by-side bar charts to compare the most frequent words by the two centuries. 
  
```{r}
df %>% 
  filter(century=='21') %>% 
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(type, word, sort = TRUE) %>% 
  head(10) %>% 
  ggplot(aes(x = n, y = reorder(word, n))) +
  geom_col() +
  labs(y = '', x = 'Frequency')
```

```{r}
df %>%
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(century, word, sort = TRUE) %>% 
  group_by(century) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, by = n, within = century)) %>%
  ggplot(aes(n, word)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~century, scales = "free") +
  labs(x = "Contribution to sentiment",
       y = NULL)+
  scale_y_reordered() 
```


**3. Word Cloud**

  a. Plot the word cloud of the words in the descriptions in the movies/TV Shows in the 20th century.
  
```{r}
library(wordcloud) 
pal <- brewer.pal(8,"Dark2")

df %>%
  filter(century=='20') %>% 
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(word, sort = TRUE) %>%
  with(wordcloud(word, n, random.order = FALSE, max.words = 50, colors=pal))
```

  b. Plot the word cloud of the words in the descriptions in the movies/TV Shows in the 21st century. 
  
```{r}
df %>%
  filter(century=='21') %>% 
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(word, sort = TRUE) %>%
  with(wordcloud(word, n, random.order = FALSE, max.words = 50, colors=pal))
```


**4. Sentiment Analysis**

  a. Is movies/TV Shows in the 21st century tends to be more positive than those in 20th century?  Use the sentiment analysis by `Bing` lexicons to address the question. 
  
```{r}
df %>%
    unnest_tokens(input = description, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(century, word, sort = TRUE) %>%
    group_by(century) %>% 
    inner_join(get_sentiments("bing")) %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    group_by(century) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(century, n, fill=sentiment))+geom_col(position = 'fill')+
    labs(y='Relative Frequency', x ='',caption='Movies/TV shows in the 21st century tend to be less positive than in the 20th century')
```

  b. Do sentiment analysis using `nrc` and `afinn` lexicons.  Give your comments on the results.

```{r}
df %>%
    unnest_tokens(input = description, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(type, word, sort = TRUE) %>%
    group_by(type) %>% 
    inner_join(get_sentiments("nrc")) %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    group_by(type) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(sentiment, n, fill=type))+geom_col(position = 'fill')+
    labs(y='Relative Frequency', x ='', caption='The frequency of most of these words is the same for both movies and TV shows.')
```


```{r}
df %>%
    unnest_tokens(input = description, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(type, word, sort = TRUE) %>%
    group_by(type) %>% 
    inner_join(get_sentiments("afinn")) %>%
    mutate(sentiment = value) %>% 
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    group_by(type) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(type, n, fill=factor(sentiment)))+geom_col(position = 'dodge')+
    labs(y='Relative Frequency', fill = 'Sentiment', x = '', caption='The distribution of sentiment is relatively the same for both movies and TV shows, 
         but Movies have a bit more negative than TV shows.')
```

**5. Modeling**

  a. Use the description to predict if a movie/TV show is in 20th or 21st century. Give the accuracy and plot the confusion matrix table. 
  
```{r}
library(caret)
library(themis)
library(textrecipes)
```

```{r}
# Select data and set target 
df1 <- df %>% 
  select(century, description) %>%
  mutate(target = century) 
```

```{r}
# Convert text data to numeric variables
a <- recipe(target~description,
       data = df1) %>% 
  step_tokenize(description) %>% 
  step_tokenfilter(description, max_tokens = 50) %>% 
  step_tfidf(description) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_smote(target) %>% 
  prep()
df1 <- juice(a)
```

```{r}
# Using Caret for modeling
set.seed(2021)
splitIndex <- createDataPartition(df1$target, p = .7, 
                                  list = FALSE)
df_train <- df1[ splitIndex,]
df_test <- df1[-splitIndex,]

forest_ranger <- train(target~., data=df_train, 
                        method = "ranger")

pred <- predict(forest_ranger, df_test)

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
d = data.frame(pred = pred, obs = df_test$target)

library(yardstick)

d %>% conf_mat(pred, obs) %>% autoplot
```

  
  b. Create variable century2 taking three following values. (Hint: You can use the case_when function to do this)

    - `21` if released_year is greater or equal to 2000
    
    - `second_half_20`if released_year is greater than or equal to 1950 and less than 2000
    
    - `first_half_20` otherwise
    
```{r}
df$century2 <- case_when(df$release_year >= 2000 ~ '21',
                         df$release_year >= 1950 ~ 'second_half_20',
                         TRUE ~ 'first_half_20')
```
    

  Predict century2 using the descriptions. Give the accuracy and plot the confusion matrix table. (Notice that the codes for 8 should still work for this question)
  
```{r}
# Select data and set target 
df2 <- df %>% 
  select(century2, description) %>%
  mutate(target = century2) 
```

```{r}
# Convert text data to numeric variables
a <- recipe(target~description,
       data = df2) %>% 
  step_tokenize(description) %>% 
  step_tokenfilter(description, max_tokens = 50) %>% 
  step_tfidf(description) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_smote(target) %>% 
  prep()
df2 <- juice(a)
```

```{r}
# Using Caret for modeling
set.seed(2021)
splitIndex <- createDataPartition(df2$target, p = .7, 
                                  list = FALSE)
df_train <- df2[ splitIndex,]
df_test <- df2[-splitIndex,]

forest_ranger <- train(target~., data=df_train, 
                        method = "ranger")

pred <- predict(forest_ranger, df_test)

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
d = data.frame(pred = pred, obs = df_test$target)

library(yardstick)

d %>% conf_mat(pred, obs) %>% autoplot
```


**6.** Create another categorical variable from the data and do the following

```{r}
df$newness <- case_when(df$release_year >= 2020 ~ 'New',
                         df$release_year >= 2017 ~ 'Old',
                         TRUE ~ 'Very Old')
```


    - Plot side-by-side word frequency by different categories of the newly created variable
    
```{r}
df %>%
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(newness, word, sort = TRUE) %>% 
  group_by(newness) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder_within(word, by = n, within = newness)) %>%
  ggplot(aes(n, word)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~newness, scales = "free") +
  labs(x = "Contribution to sentiment",
       y = NULL)+
  scale_y_reordered() 
```
    
    - Plot word clouds on different categories of the newly created variable
    
```{r}
pal <- brewer.pal(8,"Dark2")

df %>%
  filter(newness=='New') %>% 
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(word, sort = TRUE) %>%
  with(wordcloud(word, n, random.order = FALSE, max.words = 50, colors=pal))
```
    
```{r}
pal <- brewer.pal(8,"Dark2")

df %>%
  filter(newness=='Old') %>% 
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(word, sort = TRUE) %>%
  with(wordcloud(word, n, random.order = FALSE, max.words = 50, colors=pal))
```

```{r}
pal <- brewer.pal(8,"Dark2")

df %>%
  filter(newness=='Very Old') %>% 
  unnest_tokens(input = description, output = word) %>% 
  anti_join(get_stopwords()) %>% 
  count(word, sort = TRUE) %>%
  with(wordcloud(word, n, random.order = FALSE, max.words = 50, colors=pal))
```

    - Do sentiment analysis to compare different categories of the newly created variable
    
```{r}
df %>%
    unnest_tokens(input = description, output = word) %>% 
    anti_join(get_stopwords()) %>% 
    count(newness, word, sort = TRUE) %>%
    group_by(newness) %>% 
    inner_join(get_sentiments("bing")) %>%
    filter(!is.na(sentiment)) %>%
    count(sentiment, sort = TRUE) %>% 
    group_by(newness) %>% 
    mutate(n = n/sum(n)) %>% 
    ggplot(aes(newness, n, fill=sentiment))+geom_col(position = 'fill')+
    labs(y='Relative Frequency', x ='')
```
    
    - Predict the newly created variable using the description. Give the accuracy and plot the confusion matrix table. 
    
```{r}
# Select data and set target 
df3 <- df %>% 
  select(newness, description) %>%
  mutate(target = newness) 
```

```{r}
# Convert text data to numeric variables
a <- recipe(target~description,
       data = df3) %>% 
  step_tokenize(description) %>% 
  step_tokenfilter(description, max_tokens = 50) %>% 
  step_tfidf(description) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_smote(target) %>% 
  prep()
df3 <- juice(a)
```

```{r}
# Using Caret for modeling
set.seed(2021)
splitIndex <- createDataPartition(df3$target, p = .7, 
                                  list = FALSE)
df_train <- df3[ splitIndex,]
df_test <- df3[-splitIndex,]

forest_ranger <- train(target~., data=df_train, 
                        method = "ranger")

pred <- predict(forest_ranger, df_test)

cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
d = data.frame(pred = pred, obs = df_test$target)

library(yardstick)

d %>% conf_mat(pred, obs) %>% autoplot
```

