---
output:
  html_document:
    theme: united
title: "Math 421 - Midterm"
author: "Sophia Drezek"
---

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# {.tabset}

## I. Data Wranggling

1. Download the data file `hdd0318cy.sas7bdat`.  

2. Use `read_sas` in library `haven` to read the data. 


```{r, include=FALSE}
library(haven)

setwd("C:/Users/student/OneDrive/OneDrive - Bryant University/Senior Semester 1/MATH 421/Assignments/Math 421")
df <- read_sas('hdd0318cy.sas7bdat')
```

```{r, include=FALSE}
head(df)
```
    
3. Filter the data to have only patients of the year 2018 (`yod=2018`)

```{r, include=FALSE}
table(df$yod)
```

```{r, include=FALSE}
library(tidyverse)
library(dplyr)

df <- df %>% filter(yod==18)
```

    
4. Select to work with only following variables: 

```{r, include=FALSE}
df <- df %>% select("yod", "payfix","pay_ub92","age",  
                      "sex","raceethn","provider","moa", 
                      "yoa","mod","admtype", "asource" , 
                      "preopday" ,"los", "service" , "icu","ccu",    
                      "dispub92", "payer"  ,"drg","trandb", 
                      "randbg","randbs","orr", "anes","seq",   
                      "lab","dtest", "ther","blood","phar", 
                      "other","patcon","bwght","total","tot" ,  
                      "ecodub92","b_wt","pt_state","diag_adm","ancilar" ,
                      "campus","er_fee","er_chrg","er_mode","obs_chrg",
                      "obs_hour","psycchrg","nicu_day")
```


*Notice*:  You may want to save the current data to your computer for easy access later.  To save the data file use `write_csv(df, 'midterm.csv')`, for example.  

```{r, include=FALSE}
write_csv(df, 'midterm.csv')
```

```{r}
library(tidyverse)
library(dplyr)

setwd("C:/Users/student/OneDrive/OneDrive - Bryant University/Senior Semester 1/MATH 421/Assignments/Math 421")

df <- read_csv('midterm.csv')

head(df)
```

5. What are variables that have missing values?

```{r}
colSums(is.na(df))
```

6. Remove all variables with missing values. 

```{r}
df1 <- df %>% select(-'payfix',-'raceethn',-'admtype',-'asource',-'preopday',-'bwght',-'ecodub92',-'pt_state',-'diag_adm',-'er_mode', -'obs_hour',-'nicu_day')
```

```{r}
head(df1)
```


```{r}
df1 <- df1 %>% 
  mutate(sex = as.factor(sex),
         moa = as.factor(moa),
         mod = as.factor(mod),
         campus = as.factor(campus),
         provider = as.factor(provider))
```
 
7. Refer to the data description in the file `HDD2015-18cy6-20-19.docx`, which variable recording the month of admission?, which variable recording the month of discharge?

Month of admission - moa
Month of discharge - mod

8. Which month admitted the most number of patients? Which month admitted the most number of male patients?

```{r}
df1 %>% count(moa) %>%  arrange(-n)
```

```{r}
df1 %>% filter(sex==1) %>% 
  group_by(moa) %>% count(moa) %>%  arrange(-n)
```

9. Which month has the most number of teenage female patients?

```{r}
df1 %>% filter(sex==2, age>12, age<20) %>% count(moa) %>%  arrange(-n)
```

10. Which provider has the most number of female patients in October?

```{r}
df1 %>% filter(sex==2, moa==10) %>% 
  count(provider) %>%  arrange(-n)
```

11. Are female patients older than male patients, on average? 

```{r}
df1 %>% group_by(sex) %>% 
  summarize(avg_age = mean(age)) %>% 
  arrange(-avg_age)
```

12. Calculate the average age of patients by months. Which month has the oldest patients on average age?

```{r}
df1 %>% group_by(moa) %>% 
  summarize(avg_age = mean(age)) %>% 
  arrange(-avg_age)
```

13. What is the name of the provider that has the highest total charge?

```{r}
df1 %>% group_by(provider) %>% 
  summarize(max_total_charge = max(tot)) %>% 
  arrange(-max_total_charge)
```

14. What is the name of the provider that has the least total charge for teenage male on average?

```{r}
df1 %>% filter(sex==1, age>12, age<20) %>% 
  group_by(provider) %>% 
  summarize(avg_total_charge = mean(tot)) %>% 
  arrange(avg_total_charge)
```

15. Create a season (Spring, Summer, Fall, Winter) variable. Calculate the length of stays by season.  Which season has the longest length of stays on average?

```{r}
df1$season <- case_when(df1$moa %in% c(12,1,2) ~ 'Winter',
                        df1$moa %in% c(3,4,5) ~ 'Spring',
                        df1$moa %in% c(6,7,8) ~ 'Summer',
                        TRUE ~ 'Fall')
```

```{r}
df1 %>% group_by(season) %>% 
  summarize(avg_LOS = mean(los)) %>% 
  arrange(-avg_LOS)
```

16. On average, how much a 20 year-old male get charged for staying 1 day in the Fall season?

```{r}
df1 %>% filter(sex==1, age==20, season=='Fall', los==1) %>%
  summarize(avg_charge = mean(tot))
```

17. Write a paragraph to summarize the section and give your comments on the results. You could do some other calculations to support your points. 

There were a lot of interesting findings in this data. When looking at the monthly breakdown, the most patients were admitted in the month of October and the average age of both males and females admitted were around 50 years old with males being slightly higher at 51.49 years old and females being 50.86 years old. The average length of stay per season was just under 5 days. I then looked at the breakdown of total charges. Provider 7214 has the highest total charge, which is 3,402,056 dollars and the next highest charge is just under 3 million dollars. The least total charge for a teen male was approximately 10,000 dollars, which is very high for the lowest charge. Another shocking finding was that the average cost of a one day stay for a 20 year old male is 16,511 dollars. Just for one day! These were the most significant findings. 

-------

## II. Data Visualization

```{r}
head(df1)
```


Continue with the data from part I. 

1. Provides at least 10 meaningful plots. Comments on the plots. All plots should have title, caption, appropriate labels on x and y-axis


```{r}
plot1 <- df1 %>% group_by(sex) %>% count(moa) %>% 
  ggplot()+ 
  geom_col(mapping=aes(x=moa, y=n, fill=sex))+
  labs(x='Month of Admission', 
       y = 'Number of Patients',
       fill="Sex",
       title = 'Number of Patients Admitted per Month', 
       caption = 'Jan, May, Aug, and Oct have the highest intake of patients & mostly females are admitted.')

plot1
```

```{r}
ggsave(filename = 'plot1.png', plot = plot1)
```

```{r}
plot2 <- df1 %>% group_by(provider) %>% 
  summarize(avg_total_charge = mean(tot)) %>%
  ggplot()+ 
  geom_col(mapping=aes(x=provider, y=avg_total_charge))+
  labs(x='Provider', 
       y = 'Average Total Charge',
       title = 'Average Total Charge by Each Provider', 
       caption = 'Providers 7205, 7213, and 7215 have the highest average total charge.')

plot2
```
```{r}
ggsave(filename = 'plot2.png', plot = plot2)
```

```{r}
plot3 <- df1 %>% group_by(provider,sex) %>% count(provider) %>% 
  ggplot()+ 
  geom_col(mapping=aes(x=provider, y=n, fill=sex), position = 'dodge')+
  labs(x='Provider', 
       y = 'Number of Patients',
       fill="Sex",
       title = 'Number of Patients Admitted by Each Provider', 
       caption = 'Providers 7205 admits the highest number of patients.')

plot3
```

```{r}
ggsave(filename = 'plot3.png', plot = plot3)
```

```{r}
plot4 <- df1 %>% group_by(provider,sex) %>% 
  summarize(avg_los = mean(los)) %>%
  ggplot()+ 
  geom_col(mapping=aes(x=provider, y=avg_los, fill=sex),
           position="dodge")+
  labs(x='Provider', 
       y = 'Average Length of Stay',
       fill = 'Sex',
       title = 'Average Length of Stay by Each Provider', 
       caption = 'Provider 7215 has the highest average length of stay per patient for both males and females.')

plot4
```

```{r}
ggsave(filename = 'plot4.png', plot = plot4)
```

```{r}
plot5 <- df1 %>% ggplot()+ 
  geom_point(mapping=aes(x=age, y=los))+
  facet_wrap(~sex)+
  labs(x='Age', 
       y = 'Length of Stay',
       title = 'Length of Stay by Age and Sex', 
       caption = 'This shows the time each patient stayed broken down by age and sex. 
       There is no major difference between males and females length of stay.')

plot5
```

```{r}
ggsave(filename = 'plot5.png', plot = plot5)
```

```{r}
plot6 <- df %>% ggplot()+
  geom_histogram(mapping = aes(x = age))+
  labs(x='Age', 
       y = 'Number of Patients',
       title = 'Number of Patients Admitted by Age', 
       caption = 'Distribution of patients admitted by age.
       Older patients are admitted more often.')

plot6
```

```{r}
ggsave(filename = 'plot6.png', plot = plot6)
```

```{r}
plot7 <- df1 %>% ggplot()+ 
  geom_point(mapping=aes(x=tot, y=los))+
  geom_smooth(mapping=aes(x=tot, y=los))+
  labs(x='Total Cost', 
       y = 'Length of Stay',
       title = 'Total Cost vs Length of Stay', 
       caption = 'On average, the total charge increases the more days a patient is in the hospital.')

plot7
```

```{r}
ggsave(filename = 'plot7.png', plot = plot7)
```

```{r}
plot8 <- df1 %>% group_by(season) %>% 
  summarize(avg_los = mean(los)) %>%
  ggplot()+ 
  geom_col(mapping=aes(x=season, y=avg_los))+
  labs(x='Season', 
       y = 'Average Length of Stay',
       title = 'Average Length of Stay by Season', 
       caption = 'The average length of stay is approximately the same for all 4 seasons.')

plot8
```

```{r}
ggsave(filename = 'plot8.png', plot = plot8)
```

```{r}
plot9 <- df1 %>% ggplot()+
  geom_boxplot(mapping = aes(x = age, y = season))+
  labs(x='Age', 
       y = 'Season',
       title = 'Distrubtion of Age per Season', 
       caption = 'This shows the distrubtion of age per season.
       Each season has a relatively similar distribution.')

plot9
```
```{r}
ggsave(filename = 'plot9.png', plot = plot9)
```


```{r}
plot10 <- df1 %>% ggplot()+ 
  geom_point(mapping=aes(x=age, y=tot, color=sex))+
  facet_wrap(~season)+
  labs(x='Age', 
       y = 'Total Charge',
       fill='Sex',
       title = 'Age vs. Total Charge per Season', 
       caption = 'This shows the correlation between age and total charge broken down by season.
       For the most part, age does not impact the total charge.')


plot10
```
```{r}
ggsave(filename = 'plot10.png', plot = plot10)
```


2. Make an animation plot. 

```{r}
library(gganimate)
```

```{r}
plot11 <- df1 %>% group_by(provider,moa) %>% count(provider) %>% 
  ggplot()+
  geom_col(mapping=aes(x = moa, y = n))+
  transition_states(provider) +
  labs(title = 'Provider: {closest_state}')+
  labs(x='Month of Admission', 
       y = 'Number of Patients',
       title = 'Number of Patients Admitted each Month by Provider', 
       caption = 'This shows the number of patients admitted each month by provider.')

plot11
```


3. Write a paragraph to summarize the section and give your comments on the results. 

I first looked at the number of patients admitted per month. January, May, August, and October have the highest intake of patients, although all months are similar. In all months, there were more females admitted than males. I then looked at the data breakdown by provider, starting with the average total charge by each provider. Providers 7215, 7205, and 7213 have the highest average total charge, but provider 7215 is much higher than the rest. In regard to the number of patients admitted by each provider, provider 7205 admits the highest number of patients for both males and females. Provider 7215 has the highest average length of stay per patient for both males and females, which is probably why their average total cost is higher than the other providers. There was no major difference between males and females’ length of stay. When looking at the distribution of patients admitted by age, I found that older patients are admitted more often. On average, the total charge increases the more days a patient is in the hospital, which makes a lot of sense. Other findings include that the average length of stay is approximately the same for all 4 seasons and each season has a relatively similar distribution. For the most part, age does not impact the total charge.

-------

## III. Predictive Models

Continue with the data from part I. Make sure you do not have any missing values in the data. Use the follows as the target and input variables: 

```{r}
colSums(is.na(df1))
```

```{r}
head(df1)
```


*Target Variable*: Create the target variable taking value of 

  - `low` if the total charge of a patient (`tot`) is smaller than the median of the total charge, and

  - `high` otherwise. 
  
```{r}
df1$target <- case_when(df1$tot < median(df1$tot) ~ 'low',
                        TRUE ~ 'high')
```

*Input Variables*:

  - "age","sex","provider","moa","mod","campus", 'los'
  
```{r}
df2 <- df1 %>% select('age','sex','provider','moa','mod',
                      'campus', 'los', 'target')
```

-------

```{r}
head(df2)
```

1. Use `filter` function to filter out rows where `raceethn==''` or `admtype==''`. Make sure all the categorical variables are factor, numeric variables are numeric. Set Training : Testing Split = 10 : 90 


```{r}
library(caret)
splitIndex <- createDataPartition(df2$target, p = .90, 
                                  list = FALSE)
df_train <- df2[ splitIndex,]
df_test <- df2[-splitIndex,]
```

2. Train a decision tree using `rpart`.  Plot the decision tree. Plot the variable importance ranked by the tree. 

```{r}
library(rpart)

tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 5))
```

```{r}
library(rattle)
fancyRpartPlot(tree_model)
```

```{r}
tree_model$variable.importance
```

```{r}
barplot(tree_model$variable.importance)
```


3. Using caret for this question. Set `Training Control` to be: Use Cross-Validation of 5 folds across all models.  Train & tune at least 2 different models (i.e. two different values for `method=` in the train function of caret).  Plot the hyper-parameter tuning plots for each model.

```{r}
trControl = trainControl(method = "cv",
                         number = 5)

tuneGrid = expand.grid(mtry = 2:4,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(1:10))

forest_ranger <- train(target~., data=df_train, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
```

```{r}
plot(forest_ranger)
```


```{r}
df_test <- df_test %>% 
  mutate(target = as.factor(target))
```

```{r}
pred <- predict(forest_ranger, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
trControl = trainControl(method = "cv",
                         number = 5)

tuneGrid = expand.grid(predFixed = 2:5,
                       minNode = 10)

forest_Rborist <- train(target~., data=df_train, 
                    method = "Rborist", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
```

```{r}
plot(forest_Rborist)
```

```{r}
df_test <- df_test %>% 
  mutate(target = as.factor(target))
```

```{r}
pred <- predict(forest_Rborist, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

4. Plot the comparison of the models in 3. 

```{r}
results <- resamples(list('Ranger Forest' = forest_ranger,
                          'Rborist Forest' = forest_Rborist))
bwplot(results)
```

5. What is your final selection for the model? Test the accuracy of your final model on the test data. 

```{r}
pred <- predict(forest_ranger, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

6. Create another `target` variable (binary), decide the input variables and redo 1 to 5. 

```{r}
df1$target <- case_when(df1$los < median(df1$los) ~ 'short',
                        TRUE ~ 'long')
```

```{r}
df3 <- df1 %>% select('age','sex','provider','moa','mod',
                      'campus', 'tot', 'yoa', 'yod','target')
```

```{r}
library(caret)
splitIndex <- createDataPartition(df2$target, p = .90, 
                                  list = FALSE)
df_train <- df2[ splitIndex,]
df_test <- df2[-splitIndex,]
```


```{r}
library(rpart)

tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 5))
```

```{r}
library(rattle)
fancyRpartPlot(tree_model)
```

```{r}
tree_model$variable.importance
```

```{r}
barplot(tree_model$variable.importance)
```

```{r}
trControl = trainControl(method = "cv",
                         number = 5)

tuneGrid = expand.grid(mtry = 2:4,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(1:10))

forest_ranger <- train(target~., data=df_train, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
```

```{r}
plot(forest_ranger)
```

```{r}
df_test <- df_test %>% 
  mutate(target = as.factor(target))
```

```{r}
pred <- predict(forest_ranger, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
trControl = trainControl(method = "cv",
                         number = 5)

tuneGrid = expand.grid(predFixed = 2:5,
                       minNode = 10)

forest_Rborist <- train(target~., data=df_train, 
                    method = "Rborist", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
```

```{r}
plot(forest_Rborist)
```

```{r}
df_test <- df_test %>% 
  mutate(target = as.factor(target))
```

```{r}
pred <- predict(forest_Rborist, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

```{r}
results <- resamples(list('Ranger Forest' = forest_ranger,
                          'Rborist Forest' = forest_Rborist))
bwplot(results)
```

```{r}
pred <- predict(forest_ranger, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$target)
cm$overall[1]
```

7. Write a paragraph to summarize the section and give your comments on the results. 
I ran 2 different models for each target variable. The first target variable I used was the total charge of a patient and the other target variable I used was the length of stay of a patient. The two models I used for each of the target variables were ranger and Rborist random forests. In both cases, the ranger model performed better with an accuracy score of 80%. 

-------
